---
title: "Summarising differences across genotypes"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the libraries
```{r, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(ggfortify)
library(here)
library(patchwork)
```

Write a quick standard error function
```{r}
se <- function(x){
  sqrt(var(x, na.rm = TRUE)/
       sum(!is.na(x) == TRUE))
}
```

Read in the dataset, and create a dataset with only adults
```{r}
pog <- read_csv(here("data/pogona-data.csv"))

pogA <- 
  pog %>%
  filter(Age == "Adult")
```

Relevel the genotype variable so everything is compared to ZZf as that's the key comparison...

```{r}
pogA <-
  pogA %>%
  mutate(Genotype = relevel(as.factor(Genotype), ref = "ZZf"))
```


Sample composition by SVL.

Figure S1

```{r, echo = FALSE}
  ggplot(pog[!is.na(pog$Genotype), ], aes(x = SVL, 
                                   colour = Genotype, fill = Genotype)) +
  geom_histogram(alpha = 0.2, bins = 20) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  theme(legend.position = "right", legend.title = element_blank(),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(~ Genotype)

#ggsave(here("figures/samplecomposition-SVL.png"), height = 4)

```



1a. What is the range of bite forces?

```{r}
bf <-
  pog %>%
  summarise(min(BiteForce), 
            max(BiteForce))

kable(bf)
```


1b. What is the range of SVL?

```{r}
SVL <-
  pog %>%
  summarise(min(SVL), 
            max(SVL))

kable(SVL)
```


2. For adults, what are the means and SE of each variable?

This output is essentially Table 1

> This comes out as a long thin table but you can copy into Excel and quickly transpose to copy paste over into the manuscript. I have rounded to 2 decimal places.

```{r}
pogsumA <- 
  pogA %>%
  group_by(Genotype) %>%
  summarize_at(vars(Mass, SVL, HLroret, HLroqu, 
                    HWqu, HWmt, HDmt, BiteForce, BFnolever), 
               list(mean = ~ mean(., na.rm = TRUE), SE = se)) %>%
  mutate_if(is.numeric, round, digits = 2)

kable(pogsumA)

#write_csv(pogsumA, path = here("outputs/morphometrics-results.csv"))

```

3. Make figures showing the distributions of variables (BF, SVL, Mass) in relation to genotype 
(Figure 2)

```{r, warning = FALSE}
bfplot <-
  ggplot(pogA, aes(x = Genotype, y = BiteForce, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = BiteForce_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = BiteForce_mean,
                                    ymin = BiteForce_mean - BiteForce_SE,
                                    ymax = BiteForce_mean + BiteForce_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(0, 100) +
  ylab("Bite Force (N)")

bfnoleverplot <-
  ggplot(pogA, aes(x = Genotype, y = BFnolever, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = BFnolever_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = BFnolever_mean,
                                    ymin = BFnolever_mean - BFnolever_SE,
                                    ymax = BFnolever_mean + BFnolever_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(0, 150) +
  ylab("Bite Force no lever (N)")

massplot <-
  ggplot(pogA, aes(x = Genotype, y = Mass, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = Mass_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = Mass_mean,
                                    ymin = Mass_mean - Mass_SE,
                                    ymax = Mass_mean + Mass_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(100, 400) +
  ylab("Mass (g)")

svlplot <-
  ggplot(pogA, aes(x = Genotype, y = SVL, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = SVL_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = SVL_mean,
                                    ymin = SVL_mean - SVL_SE,
                                    ymax = SVL_mean + SVL_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(100, 250) +
  ylab("SVL (mm)")

(bfplot + bfnoleverplot) / (massplot + svlplot)

#ggsave(here("figures/Figure2-colour.png"), height = 5)
```

Black and white version


```{r, warning = FALSE}
bfplot <-
  ggplot(pogA, aes(x = Genotype, y = BiteForce, 
                colour = Genotype, fill = Genotype,  shape = Genotype)) +
  geom_jitter(alpha = 0.6, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("grey50", "grey50", "grey50")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = BiteForce_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = BiteForce_mean,
                                    ymin = BiteForce_mean - BiteForce_SE,
                                    ymax = BiteForce_mean + BiteForce_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(panel.grid.minor = element_blank(), legend.position = "none") +
  ylim(0, 100) +
  ylab("Bite Force (N)")

bfnoleverplot <-
  ggplot(pogA, aes(x = Genotype, y = BFnolever, 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_jitter(alpha = 0.6, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("grey50", "grey50", "grey50")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +

  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = BFnolever_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = BFnolever_mean,
                                    ymin = BFnolever_mean - BFnolever_SE,
                                    ymax = BFnolever_mean + BFnolever_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(panel.grid.minor = element_blank(), legend.position = "none") +
  ylim(0, 150) +
  ylab("Bite Force no lever (N)")

massplot <-
  ggplot(pogA, aes(x = Genotype, y = Mass, 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_jitter(alpha = 0.6, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("grey50", "grey50", "grey50")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +

  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = Mass_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = Mass_mean,
                                    ymin = Mass_mean - Mass_SE,
                                    ymax = Mass_mean + Mass_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(panel.grid.minor = element_blank(), legend.position = "none") +
  ylim(100, 400) +
  ylab("Mass (g)")

svlplot <-
  ggplot(pogA, aes(x = Genotype, y = SVL, 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_jitter(alpha = 0.6, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("grey50", "grey50", "grey50")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +

  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = SVL_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = SVL_mean,
                                    ymin = SVL_mean - SVL_SE,
                                    ymax = SVL_mean + SVL_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(panel.grid.minor = element_blank(), legend.position = "none") +
  ylim(100, 250) +
  ylab("SVL (mm)")

(bfplot + bfnoleverplot) / (massplot + svlplot)

#ggsave(here("figures/Figure2-bw-900dpi.png"), height = 5, dpi=900)
```


4. And for the head dimensions (Figure 3)

Colour version

```{r, warning = FALSE}
HLroretplot <-
  ggplot(pogA, aes(x = Genotype, y = HLroret, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = HLroret_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = HLroret_mean,
                                    ymin = HLroret_mean - HLroret_SE,
                                    ymax = HLroret_mean + HLroret_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(40, 70) +
  ylab("HLroret (mm)")

HLroquplot <-
  ggplot(pogA, aes(x = Genotype, y = HLroqu, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = HLroqu_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = HLroqu_mean,
                                    ymin = HLroqu_mean - HLroqu_SE,
                                    ymax = HLroqu_mean + HLroqu_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(30, 60) +
  ylab("HLroqu (mm)")

HWquplot <-
  ggplot(pogA, aes(x = Genotype, y = HWqu, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = HWqu_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = HWqu_mean,
                                    ymin = HWqu_mean - HWqu_SE,
                                    ymax = HWqu_mean + HWqu_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(30, 50) +
  ylab("HWqu (mm)")

HWmtplot <-
  ggplot(pogA, aes(x = Genotype, y = HWmt, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = HWmt_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = HWmt_mean,
                                    ymin = HWmt_mean - HWmt_SE,
                                    ymax = HWmt_mean + HWmt_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(20, 50) +
  ylab("HWmt (mm)")

HDmtplot <-
  ggplot(pogA, aes(x = Genotype, y = HDmt, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = HDmt_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = HDmt_mean,
                                    ymin = HDmt_mean - HDmt_SE,
                                    ymax = HDmt_mean + HDmt_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(20, 35) +
  ylab("HDmt (mm)")

z <- gridExtra::grid.arrange(HLroretplot, HLroquplot,
                        HWquplot, HWmtplot,
                        HDmtplot, nrow = 3)

#ggsave(file = here("figures/Figure3-colour.png"), z, height = 7.5)
```

Black and white version

```{r, warning = FALSE}
HLroretplot <-
  ggplot(pogA, aes(x = Genotype, y = HLroret, 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("grey50", "grey50", "grey50")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = HLroret_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = HLroret_mean,
                                    ymin = HLroret_mean - HLroret_SE,
                                    ymax = HLroret_mean + HLroret_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(panel.grid.minor = element_blank(), legend.position = "none") +
  ylim(40, 70) +
  ylab("HLroret (mm)")

HLroquplot <-
  ggplot(pogA, aes(x = Genotype, y = HLroqu, 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("grey50", "grey50", "grey50")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = HLroqu_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = HLroqu_mean,
                                    ymin = HLroqu_mean - HLroqu_SE,
                                    ymax = HLroqu_mean + HLroqu_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(panel.grid.minor = element_blank(), legend.position = "none") +
  ylim(30, 60) +
  ylab("HLroqu (mm)")

HWquplot <-
  ggplot(pogA, aes(x = Genotype, y = HWqu, 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("grey50", "grey50", "grey50")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = HWqu_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = HWqu_mean,
                                    ymin = HWqu_mean - HWqu_SE,
                                    ymax = HWqu_mean + HWqu_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(panel.grid.minor = element_blank(), legend.position = "none") +
  ylim(30, 50) +
  ylab("HWqu (mm)")

HWmtplot <-
  ggplot(pogA, aes(x = Genotype, y = HWmt, 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("grey50", "grey50", "grey50")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = HWmt_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = HWmt_mean,
                                    ymin = HWmt_mean - HWmt_SE,
                                    ymax = HWmt_mean + HWmt_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(panel.grid.minor = element_blank(), legend.position = "none") +
  ylim(20, 50) +
  ylab("HWmt (mm)")

HDmtplot <-
  ggplot(pogA, aes(x = Genotype, y = HDmt, 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("grey50", "grey50", "grey50")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +
  # Add mean and se from pogsumA
  geom_point(data = pogsumA, aes(y = HDmt_mean), size = 2) +
  geom_errorbar(data = pogsumA, aes(y = HDmt_mean,
                                    ymin = HDmt_mean - HDmt_SE,
                                    ymax = HDmt_mean + HDmt_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(panel.grid.minor = element_blank(), legend.position = "none") +
  ylim(20, 35) +
  ylab("HDmt (mm)")

x <- gridExtra::grid.arrange(HLroretplot, HLroquplot,
                        HWquplot, HWmtplot,
                        HDmtplot, nrow = 3)

#ggsave(file = here("figures/Figure3-bw-900dpi.png"), x, height = 7.5, dpi=900)
```


