---
title: "ANOVAs for different genotypes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the libraries
```{r, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(ggfortify)
library(here)
```

Write a quick standard error function
```{r}
se <- function(x){
  sqrt(var(x)/length(x))
}
```

Read in the dataset, and create a dataset with only adults
```{r}
pog <- read_csv(here("data/pogona-data.csv"))

pogA <- 
  pog %>%
  filter(Age == "Adult")
```

Relevel the genotype variable so everything is compared to ZZf as that's the key comparison...

```{r}
pogA <-
  pogA %>%
  mutate(Genotype = relevel(as.factor(Genotype), ref = "ZZf"))
```

Fit ANOVAs to look at differences in size, head shape and bite force in the three genotypes.

First to make the models easier to interpret, relevel the genotype variable so everything is compared to ZZf as that's the key comparison...

Next set up an empty output file
```{r}
output <- data.frame(array(dim = c(11, 15)))
names(output) <- c("response", "df1", "df2", "F", "p", 
                  "mean_ZZf", "se_ZZf",
                  "meandiff_ZWf", "se_ZWf", "t_ZWf", "p_ZWf",
                  "meandiff_ZZm", "se_ZZm", "t_ZZm", "p_ZZm")

# Short function to help write output files
write_output <- function(model, response, n){
  
  x <- tidy(anova(model))
  y <- summary(model)
    
  output[n, "response"] <- response
  output[n, "df1"] <- x$df[1]
    output[n, "df2"] <- x$df[2]
    output[n, "F"] <- x$statistic[1]
    output[n, "p"] <- x$p.value[1]
    output[n, "mean_ZZf"] <- y$coefficients[1]
    output[n, "se_ZZf"] <- y$coefficients[4]
    output[n, "meandiff_ZWf"] <- y$coefficients[2]
    output[n, "se_ZWf"] <- y$coefficients[5]
    output[n, "t_ZWf"] <- y$coefficients[8]
    output[n, "p_ZWf"] <- y$coefficients[11]
    output[n, "meandiff_ZZm"] <- y$coefficients[3]
    output[n, "se_ZZm"] <- y$coefficients[5]
    output[n, "t_ZZm"] <- y$coefficients[9]
    output[n, "p_ZZm"] <- y$coefficients[12]
    
    return(output)
}
```

Mass
```{r}
# Fit the model
m1 <- lm(log(Mass) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m1)
tidy(anova(m1))
summary(m1)

output <- write_output(m1, "log(Mass)", 1)
```

Snout-vent-length
(SVL)
```{r}
# Fit the model
m2 <- lm(log(SVL) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m2)
tidy(anova(m2))
summary(m2)

output <- write_output(m2, "log(SVL)", 2)
```

Bite Force (with a lever correction). Greatest value from four trials (B4 in original spreadsheet). 
```{r}
# Fit the model
m3 <- lm(log(BiteForce) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m3)
tidy(anova(m3))
tidy(summary(m3))

output <- write_output(m3, "log(Bite Force)", 3)
```

Bite Force without a lever correction
```{r}
# Fit the model
m4 <- lm(log(BFnolever) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m4)
tidy(anova(m4))
tidy(summary(m4))

output <- write_output(m4, "log(Bite Force nl)", 4)
```

Head length directly from the tip of the rostrum to the posterior end of the retroarticular process 
(HLroret and later HLrr)
```{r}
# Fit the model
m5 <- lm(log(HLroret) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m5)
tidy(anova(m5))
tidy(summary(m5))

output <- write_output(m5, "log(HLroret)", 5)
```

Head length from the tip of the rostrum to the posteroventral corner of the tympanum (approximately the position of the quadrate-articular jaw joint) (HLrq), 
(HLroqu and in the manuscript HLrq)
```{r}
# Fit the model
m6 <- lm(log(HLroqu) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m6)
tidy(anova(m6))
tidy(summary(m6))

output <- write_output(m6, "log(HLroqu)", 6)
```

Head width at the base of the tympanum 
(HWqu)
```{r}
# Fit the model
m7 <- lm(log(HWqu) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m7)
tidy(anova(m7))
tidy(summary(m7))

output <- write_output(m7, "log(HWqu)", 7)
```

Head width at the midpoint of the temporal region 
(HWmt)
```{r}
# Fit the model
m8 <- lm(log(HWmt) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m8)
tidy(anova(m8))
tidy(summary(m8))

output <- write_output(m8, "log(HWmt)", 8)
```

Head depth at the midpoint of the temporal region 
(HDmt)
```{r}
# Fit the model
m9 <- lm(log(HDmt) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m9)
tidy(anova(m9))
tidy(summary(m9))

output <- write_output(m9, "log(HDmt)", 9)
```

Temperature measured using a temperature gun between trial 1 and trial 2
(Temp)
```{r}
# Fit the model
m10 <- lm(log(Temp) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m10)
tidy(anova(m10))
summary(m10)

output <- write_output(m10, "log(Temp)", 10)
```

Egg incubation temperature
(IncTemp)
```{r}
# Fit the model
m11 <- lm(log(IncTemp) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m11)
tidy(anova(m11))
summary(m11)

output <- write_output(m11, "log(IncTemp)", 11)
```


Write all outputs to file
```{r}
write_csv(output, path = here("outputs/genotype-anova-results.csv"))
```


Plots of log raw head dimensions against SVL.

SVL vs HLroret

```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(HLroret), 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(HLroret) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/svl-HLroret.png"))
```


SVL vs HLroqu

```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(HLroqu), 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(HLroqu) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/svl-HLroqu.png"))
```

SVL vs HWmt

```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(HWmt), 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(HWmt) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/svl-Hwqu.png"))
```


SVL vs HWqu

```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(HWqu), 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(HWqu) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/svl-Hwqu.png"))
```

SVL vs HDmt

```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(HDmt), 
                colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(HDmt) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/svl-HDmt.png"))
```

