---
title: "Allometry"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the libraries
```{r, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(ggfortify)
library(here)
```

Read in the dataset, relevel the genotype variable so everything is compared to ZZf as that's the key comparison, and create a dataset for each genotype.

```{r}
pog <- read_csv(here("data/pogona-data.csv"))

pog <-
  pog %>%
  mutate(Genotype = relevel(as.factor(Genotype), ref = "ZZf"))

pogF <-
  pog %>%
  filter(Genotype == "ZWf")

pogM <-
  pog %>%
  filter(Genotype == "ZZm")

pogSRF <-
  pog %>%
  filter(Genotype == "ZZf")
```

Now fit the models and save the outputs
Remember to use log 10 not ln!!!


Forms part of Table 3.

```{r}
modelF <- lm(log10(BiteForce) ~ log10(SVL), data = pogF)
tidy(anova(modelF))
summary(modelF)

modelM <- lm(log10(BiteForce) ~ log10(SVL), data = pogM)
tidy(anova(modelM))
summary(modelM)

modelSRF <- lm(log10(BiteForce) ~ log10(SVL), data = pogSRF)
anova(modelSRF)
summary(modelSRF)

output <- data.frame(array(dim = c(3, 7)))
names(output) <- c("genotype", "slope", "SE", "df", "t", "p", "adj.r2")

output["genotype"] <- c("ZWf", "ZZm", "ZZf")
output["slope"] <- c(summary(modelF)$coefficients[2],
                     summary(modelM)$coefficients[2],
                     summary(modelSRF)$coefficients[2])
output["SE"] <- c(summary(modelF)$coefficients[4],
                     summary(modelM)$coefficients[4],
                     summary(modelSRF)$coefficients[4])
output["df"] <- c(summary(modelF)$df[2],
                     summary(modelM)$df[2],
                     summary(modelSRF)$df[2])
output["t"] <- c(summary(modelF)$coefficients[6],
                     summary(modelM)$coefficients[6],
                     summary(modelSRF)$coefficients[6])
output["p"] <- c(summary(modelF)$coefficients[8],
                     summary(modelM)$coefficients[8],
                     summary(modelSRF)$coefficients[8])
output["adj.r2"] <- c(summary(modelF)$adj.r.squared,
                      summary(modelM)$adj.r.squared,
                      summary(modelSRF)$adj.r.squared)

#write_csv(output, path = here("outputs/allometry-results-SVL.csv"))
```


Now fit the models and save the outputs
Remember to use log 10 not ln!!!

Forms part of Table 3.

```{r}
modelF <- lm(log10(BiteForce) ~ log10(Mass), data = pogF)
tidy(anova(modelF))
summary(modelF)

modelM <- lm(log10(BiteForce) ~ log10(Mass), data = pogM)
tidy(anova(modelM))
summary(modelM)

modelSRF <- lm(log10(BiteForce) ~ log10(Mass), data = pogSRF)
anova(modelSRF)
summary(modelSRF)

output <- data.frame(array(dim = c(3, 7)))
names(output) <- c("genotype", "slope", "SE", "df", "t", "p", "adj.r2")

output["genotype"] <- c("ZWf", "ZZm", "ZZf")
output["slope"] <- c(summary(modelF)$coefficients[2],
                     summary(modelM)$coefficients[2],
                     summary(modelSRF)$coefficients[2])
output["SE"] <- c(summary(modelF)$coefficients[4],
                     summary(modelM)$coefficients[4],
                     summary(modelSRF)$coefficients[4])
output["df"] <- c(summary(modelF)$df[2],
                     summary(modelM)$df[2],
                     summary(modelSRF)$df[2])
output["t"] <- c(summary(modelF)$coefficients[6],
                     summary(modelM)$coefficients[6],
                     summary(modelSRF)$coefficients[6])
output["p"] <- c(summary(modelF)$coefficients[8],
                     summary(modelM)$coefficients[8],
                     summary(modelSRF)$coefficients[8])
output["adj.r2"] <- c(summary(modelF)$adj.r.squared,
                      summary(modelM)$adj.r.squared,
                      summary(modelSRF)$adj.r.squared)

#write_csv(output, path = here("outputs/allometry-results-Mass.csv"))
```


Now fit the models and save the outputs
Remember to use log 10 not ln!!!

Forms part of Table in SI

HLroret

```{r}
modelF <- lm(log10(BiteForce) ~ log10(HLroret), data = pogF)
tidy(anova(modelF))
summary(modelF)

modelM <- lm(log10(BiteForce) ~ log10(HLroret), data = pogM)
tidy(anova(modelM))
summary(modelM)

modelSRF <- lm(log10(BiteForce) ~ log10(HLroret), data = pogSRF)
anova(modelSRF)
summary(modelSRF)

output <- data.frame(array(dim = c(3, 7)))
names(output) <- c("genotype", "slope", "SE", "df", "t", "p", "adj.r2")

output["genotype"] <- c("ZWf", "ZZm", "ZZf")
output["slope"] <- c(summary(modelF)$coefficients[2],
                     summary(modelM)$coefficients[2],
                     summary(modelSRF)$coefficients[2])
output["SE"] <- c(summary(modelF)$coefficients[4],
                     summary(modelM)$coefficients[4],
                     summary(modelSRF)$coefficients[4])
output["df"] <- c(summary(modelF)$df[2],
                     summary(modelM)$df[2],
                     summary(modelSRF)$df[2])
output["t"] <- c(summary(modelF)$coefficients[6],
                     summary(modelM)$coefficients[6],
                     summary(modelSRF)$coefficients[6])
output["p"] <- c(summary(modelF)$coefficients[8],
                     summary(modelM)$coefficients[8],
                     summary(modelSRF)$coefficients[8])
output["adj.r2"] <- c(summary(modelF)$adj.r.squared,
                      summary(modelM)$adj.r.squared,
                      summary(modelSRF)$adj.r.squared)

#write_csv(output, path = here("outputs/allometry-results-HLroret.csv"))
```

HLroqu

```{r}
modelF <- lm(log10(BiteForce) ~ log10(HLroqu), data = pogF)
tidy(anova(modelF))
summary(modelF)

modelM <- lm(log10(BiteForce) ~ log10(HLroqu), data = pogM)
tidy(anova(modelM))
summary(modelM)

modelSRF <- lm(log10(BiteForce) ~ log10(HLroqu), data = pogSRF)
anova(modelSRF)
summary(modelSRF)

output <- data.frame(array(dim = c(3, 7)))
names(output) <- c("genotype", "slope", "SE", "df", "t", "p", "adj.r2")

output["genotype"] <- c("ZWf", "ZZm", "ZZf")
output["slope"] <- c(summary(modelF)$coefficients[2],
                     summary(modelM)$coefficients[2],
                     summary(modelSRF)$coefficients[2])
output["SE"] <- c(summary(modelF)$coefficients[4],
                     summary(modelM)$coefficients[4],
                     summary(modelSRF)$coefficients[4])
output["df"] <- c(summary(modelF)$df[2],
                     summary(modelM)$df[2],
                     summary(modelSRF)$df[2])
output["t"] <- c(summary(modelF)$coefficients[6],
                     summary(modelM)$coefficients[6],
                     summary(modelSRF)$coefficients[6])
output["p"] <- c(summary(modelF)$coefficients[8],
                     summary(modelM)$coefficients[8],
                     summary(modelSRF)$coefficients[8])
output["adj.r2"] <- c(summary(modelF)$adj.r.squared,
                      summary(modelM)$adj.r.squared,
                      summary(modelSRF)$adj.r.squared)

#write_csv(output, path = here("outputs/allometry-results-HLroqu.csv"))
```

HWqu

```{r}
modelF <- lm(log10(BiteForce) ~ log10(HWqu), data = pogF)
tidy(anova(modelF))
summary(modelF)

modelM <- lm(log10(BiteForce) ~ log10(HWqu), data = pogM)
tidy(anova(modelM))
summary(modelM)

modelSRF <- lm(log10(BiteForce) ~ log10(HWqu), data = pogSRF)
anova(modelSRF)
summary(modelSRF)

output <- data.frame(array(dim = c(3, 7)))
names(output) <- c("genotype", "slope", "SE", "df", "t", "p", "adj.r2")

output["genotype"] <- c("ZWf", "ZZm", "ZZf")
output["slope"] <- c(summary(modelF)$coefficients[2],
                     summary(modelM)$coefficients[2],
                     summary(modelSRF)$coefficients[2])
output["SE"] <- c(summary(modelF)$coefficients[4],
                     summary(modelM)$coefficients[4],
                     summary(modelSRF)$coefficients[4])
output["df"] <- c(summary(modelF)$df[2],
                     summary(modelM)$df[2],
                     summary(modelSRF)$df[2])
output["t"] <- c(summary(modelF)$coefficients[6],
                     summary(modelM)$coefficients[6],
                     summary(modelSRF)$coefficients[6])
output["p"] <- c(summary(modelF)$coefficients[8],
                     summary(modelM)$coefficients[8],
                     summary(modelSRF)$coefficients[8])
output["adj.r2"] <- c(summary(modelF)$adj.r.squared,
                      summary(modelM)$adj.r.squared,
                      summary(modelSRF)$adj.r.squared)

#write_csv(output, path = here("outputs/allometry-results-HWqu.csv"))
```


HWmt

```{r}
modelF <- lm(log10(BiteForce) ~ log10(HWmt), data = pogF)
tidy(anova(modelF))
summary(modelF)

modelM <- lm(log10(BiteForce) ~ log10(HWmt), data = pogM)
tidy(anova(modelM))
summary(modelM)

modelSRF <- lm(log10(BiteForce) ~ log10(HWmt), data = pogSRF)
anova(modelSRF)
summary(modelSRF)

output <- data.frame(array(dim = c(3, 7)))
names(output) <- c("genotype", "slope", "SE", "df", "t", "p", "adj.r2")

output["genotype"] <- c("ZWf", "ZZm", "ZZf")
output["slope"] <- c(summary(modelF)$coefficients[2],
                     summary(modelM)$coefficients[2],
                     summary(modelSRF)$coefficients[2])
output["SE"] <- c(summary(modelF)$coefficients[4],
                     summary(modelM)$coefficients[4],
                     summary(modelSRF)$coefficients[4])
output["df"] <- c(summary(modelF)$df[2],
                     summary(modelM)$df[2],
                     summary(modelSRF)$df[2])
output["t"] <- c(summary(modelF)$coefficients[6],
                     summary(modelM)$coefficients[6],
                     summary(modelSRF)$coefficients[6])
output["p"] <- c(summary(modelF)$coefficients[8],
                     summary(modelM)$coefficients[8],
                     summary(modelSRF)$coefficients[8])
output["adj.r2"] <- c(summary(modelF)$adj.r.squared,
                      summary(modelM)$adj.r.squared,
                      summary(modelSRF)$adj.r.squared)

#write_csv(output, path = here("outputs/allometry-results-HWmt.csv"))
```

HDmt

```{r}
modelF <- lm(log10(BiteForce) ~ log10(HDmt), data = pogF)
tidy(anova(modelF))
summary(modelF)

modelM <- lm(log10(BiteForce) ~ log10(HDmt), data = pogM)
tidy(anova(modelM))
summary(modelM)

modelSRF <- lm(log10(BiteForce) ~ log10(HDmt), data = pogSRF)
anova(modelSRF)
summary(modelSRF)

output <- data.frame(array(dim = c(3, 7)))
names(output) <- c("genotype", "slope", "SE", "df", "t", "p", "adj.r2")

output["genotype"] <- c("ZWf", "ZZm", "ZZf")
output["slope"] <- c(summary(modelF)$coefficients[2],
                     summary(modelM)$coefficients[2],
                     summary(modelSRF)$coefficients[2])
output["SE"] <- c(summary(modelF)$coefficients[4],
                     summary(modelM)$coefficients[4],
                     summary(modelSRF)$coefficients[4])
output["df"] <- c(summary(modelF)$df[2],
                     summary(modelM)$df[2],
                     summary(modelSRF)$df[2])
output["t"] <- c(summary(modelF)$coefficients[6],
                     summary(modelM)$coefficients[6],
                     summary(modelSRF)$coefficients[6])
output["p"] <- c(summary(modelF)$coefficients[8],
                     summary(modelM)$coefficients[8],
                     summary(modelSRF)$coefficients[8])
output["adj.r2"] <- c(summary(modelF)$adj.r.squared,
                      summary(modelM)$adj.r.squared,
                      summary(modelSRF)$adj.r.squared)

#write_csv(output, path = here("outputs/allometry-results-HDmt.csv"))
```