---
title: "Bite force versus size analyses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the libraries
```{r, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(ggfortify)
library(here)
library(patchwork)
```

Read in the dataset, relevel the genotype variable so everything is compared to ZZf as that's the key comparison.

```{r}
pog <- read_csv(here("data/pogona-data.csv"))

pog <-
  pog %>%
  mutate(Genotype = relevel(as.factor(Genotype), ref = "ZZf"))
```

This creates a plot of log(BF) vs log(SVL) Figure 4a

```{r, message = FALSE}
# All individuals
log_plot<-
  ggplot(pog, aes(x = log(SVL), y = log(BiteForce), 
                colour = Genotype, shape = Genotype, fill = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(Bite Force) (N)") +
  xlab("log(SVL) (mm)") +
  labs(tag = "(a)") +
  theme(legend.position = "none") +
  theme(plot.tag = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 12))

log_plot
```

This creates a non logged plot of BF vs SVL (Figure 4b)

```{r, message = FALSE}
# All individuals
nl_plot <- 
ggplot(pog, aes(x = SVL, y = BiteForce, 
                colour = Genotype, shape = Genotype, fill = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 100)+
  xlim(50, 250) +
  ylab("Bite Force (N)") +
  xlab("SVL (mm)") +
  labs(tag = "(b)") +
  theme(plot.tag = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 12))

nl_plot

# combine a and b
log_plot / nl_plot

#ggsave(here("figures/Figure4-colour.png"), height = 8 , width = 7)
```

This creates a plot of log(BF) vs log(SVL)
(Figure 4a in grayscale without the lines)

```{r, message = FALSE}
log_plot_bw <-
ggplot(pog, aes(x = log(SVL), y = log(BiteForce), 
                colour = Genotype, fill = Genotype, shape = Genotype, size=Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2, size = 0.5) +
  scale_colour_manual(values = c("grey30", "grey30", "black")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +
  scale_size_manual(values=c(2,2,2))+
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(0, 5)+
  xlim(4, 5.5) +
  labs(tag = "(a)") +
  ylab("log(Bite Force) (N)") +
  xlab("log(SVL) (mm)")

log_plot_bw
```

This creates a plot of BF vs SVL
(Figure 4b in grayscale)

```{r, message = FALSE}
nl_plot_bw <-
 ggplot(pog, aes(x = (SVL), y = (BiteForce), 
                colour = Genotype, fill = Genotype, shape = Genotype, size=Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  #geom_smooth(method = lm, se = TRUE, alpha = 0.15, size = 0.4) +
  scale_colour_manual(values = c("grey50", "grey50", "grey50")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("grey30", "grey70", "black")) +
  scale_size_manual(values=c(2,2,2))+
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 100) +
  xlim(50, 250) +
  labs(tag = "(b)") +
  ylab("Bite Force (N)") +
  xlab("SVL (mm)") +
  theme(plot.tag = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 12))



# Combine the two
log_plot_bw / nl_plot_bw

#ggsave(here("figures/Figure4-bw-900dpi.png"), height = 8 , width = 7, dpi=900)

```




This creates a plot of log(BF) vs log(Mass) Figure S 5

```{r, message = FALSE}
# All individuals
log_plot_mass<-
  ggplot(pog, aes(x = log(Mass), y = log(BiteForce), 
                colour = Genotype, shape = Genotype, fill = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(1, 5)+
  xlim(1, 6) +
  ylab("log(Bite Force) (N)") +
  xlab("log(Mass) (g)") +
  labs(tag = "(a)") +
  theme(legend.position = "none") +
  theme(plot.tag = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 12))

log_plot_mass

#ggsave(here("figures/FigS4-col-BFlogmass.png"), height = 6 , width = 7)
```


log(SVL) and log(bite force no lever) Figure S 6


```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(BFnolever), 
                colour = Genotype, fill = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(Bite Force no lever) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/svl-bfnolever.png"))
```



Next set up an empty output file
```{r}
output <- data.frame(array(dim = c(14, 20)))
names(output) <- c("response", "predictor", "df", "F_pred", "p_pred", "F_gen", "p_gen", "F_int", "p_int", "slope_ZZf", "slope_se_ZZf", "slopediff_ZWf", "se_ZWf", "t_ZWf", "p_ZWf", "slopediff_ZZm", "se_ZZm", "t_ZZm", "p_ZZm", "adj.r2")

# Short function to help write output files
write_output <- function(model, response, predictor, n){
  
  x <- tidy(anova(model))
  y <- summary(model)
    
  output[n, "response"] <- response
  output[n, "predictor"] <- predictor
  output[n, "df"] <- x$df[2]
  output[n, "F_pred"] <- x$statistic[1]
  output[n, "p_pred"] <- x$p.value[1]
  output[n, "F_gen"] <- y$coefficients[1]
  output[n, "p_gen"] <- y$coefficients[4]
  output[n, "F_int"] <- y$coefficients[2]
  output[n, "p_int"] <- y$coefficients[5]
  output[n, "slope_ZZf"] <- y$coefficients[8]
  output[n, "slope_se_ZZf"] <- y$coefficients[11]
  output[n, "slopedifff_ZWf"] <- y$coefficients[3]
  output[n, "se_ZWf"] <- y$coefficients[5]
  output[n, "t_ZWf"] <- y$coefficients[9]
  output[n, "p_ZWf"] <- y$coefficients[12]
  output[n, "slopedifff_ZZm"] <- y$coefficients[3]
  output[n, "se_ZZm"] <- y$coefficients[5]
  output[n, "t_ZZm"] <- y$coefficients[9]
  output[n, "p_ZZm"] <- y$coefficients[12] 
  return(output)
}
```


~~ ANOVAs ~~

Mass
```{r}
# Fit the model
m1 <- lm(log(BiteForce) ~ log(Mass) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m1)
tidy(anova(m1))
summary(m1)

output <- write_output(m1, "log(Bite Force)", "log(Mass)", 1)
```

SVL
```{r}
# Fit the model
m2 <- lm(log(BiteForce) ~ log(SVL) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m2)
tidy(anova(m2))
summary(m2)

output <- write_output(m2, "log(Bite Force)", "log(SVL)", 2)
```

HLroret
```{r}
# Fit the model
m3 <- lm(log(BiteForce) ~ log(HLroret) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m3)
tidy(anova(m3))
tidy(summary(m3))

output <- write_output(m3, "log(Bite Force)", "log(HLroret)", 3)
```

HLroqu
```{r}
# Fit the model
m4 <- lm(log(BiteForce) ~ log(HLroqu) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m4)
tidy(anova(m4))
tidy(summary(m4))

output <- write_output(m4, "log(Bite Force)", "log(HLroqu)", 4)
```

HWqu
```{r}
# Fit the model
m5 <- lm(log(BiteForce) ~ log(HWqu) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m5)
tidy(anova(m5))
tidy(summary(m5))

output <- write_output(m5, "log(Bite Force)", "log(HWqu)", 5)
```

HWmt
```{r}
# Fit the model
m6 <- lm(log(BiteForce) ~ log(HWmt) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m6)
tidy(anova(m6))
tidy(summary(m6))

output <- write_output(m6, "log(Bite Force)", "log(HWmt)", 6)
```

HDmt
```{r}
# Fit the model
m7 <- lm(log(BiteForce) ~ log(HDmt) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m7)
tidy(anova(m7))
tidy(summary(m7))

output <- write_output(m7, "log(Bite Force)", "log(HDmt)", 7)
```


And get rid of the lever...

Mass

```{r}
# Fit the model
m8 <- lm(log(BFnolever) ~ log(Mass) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m8)
tidy(anova(m8))
summary(m8)

output <- write_output(m8, "log(Bite Force nl)", "log(Mass)", 8)
```

SVL
```{r}
# Fit the model
m9 <- lm(log(BFnolever) ~ log(SVL) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m9)
tidy(anova(m9))
summary(m9)

output <- write_output(m9, "log(Bite Force nl)", "log(SVL)", 9)
```

HLroret
```{r}
# Fit the model
m10 <- lm(log(BFnolever) ~ log(HLroret) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m10)
tidy(anova(m10))
tidy(summary(m10))

output <- write_output(m10, "log(Bite Force nl)", "log(HLroret)", 10)
```

HLroqu
```{r}
# Fit the model
m11 <- lm(log(BFnolever) ~ log(HLroqu) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m11)
tidy(anova(m11))
tidy(summary(m11))

output <- write_output(m11, "log(Bite Force nl)", "log(HLroqu)", 11)
```

HWqu
```{r}
# Fit the model
m12 <- lm(log(BFnolever) ~ log(HWqu) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m12)
tidy(anova(m12))
tidy(summary(m12))

output <- write_output(m12, "log(Bite Force nl)", "log(HWqu)", 12)
```

HWmt
```{r}
# Fit the model
m13 <- lm(log(BFnolever) ~ log(HWmt) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m13)
tidy(anova(m13))
tidy(summary(m13))

output <- write_output(m13, "log(Bite Force nl)", "log(HWmt)", 13)
```

HDmt
```{r}
# Fit the model
m14 <- lm(log(BFnolever) ~ log(HDmt) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m14)
tidy(anova(m14))
tidy(summary(m14))

output <- write_output(m14, "log(Bite Force nl)", "log(HDmt)", 14)
```

Write all outputs to file

Table 4

```{r}
write_csv(output, path = here("outputs/bite-force-scaling-results.csv"))
```