---
title: "PCA plot"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse)
library(knitr)
library(here)
library(ggpubr) # for fancy multi-panel plots

#need to install from GitHub 
#install.packages("devtools")
#install_github("cmartin/ggConvexHull")
#install_github("ggConvexHull")
#devtools::install_github("cmartin/ggConvexHull")
# need to install from GitHub

library(ggConvexHull)  

library(ggfortify)
library(car)
library(broom)

# Helper functions for plotting
remove_y <- 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

remove_x <-   
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

Read in the dataset, relevel the genotype variable so everything is compared to ZZf as that's the key comparison.

```{r}
pog <- read_csv(here("data/pogona-data.csv"))

pog <-
  pog %>%
  mutate(Genotype = relevel(as.factor(Genotype), ref = "ZZF"))
```

PCA for all animals (including juveniles) (n = 91).

```{r, echo = FALSE, warning = FALSE, message = FALSE}
#First process the data so we can use it...

# Perform a principal components analysis on log shape variables
pca_head <- princomp(as.matrix(select(pog, LSR_HLroret:LSR_HDmt)),
                     scores = TRUE, cor = TRUE)

# Recombine this with the full dataset
# First make a dataset of just the PC scores and label the rows with ID
# taken from the dataset as these are in the same order
pca_head_ds <- data.frame(pca_head$scores, ID = pog$ID)

# Next join this on "ID" to the original data
pog2 <- left_join(pog, pca_head_ds, by = "ID")
```

PCA outputs for all animals (including juveniles) (n = 91).
```{r}
loadings(pca_head)
summary(pca_head)
```

Loadings diagram for full sample (n = 91)

```{r}
#princomp(pca_head, cor = TRUE) 

mydata <- data.frame(unclass(loadings(pca_head)))
mydata <- mutate(mydata, variable = rownames(mydata))


library(grid)

pc1 <- 
ggplot(mydata, aes(x = Comp.1, y = variable)) +
  xlim(-0.75, 0.75) +
  geom_segment(aes(x = 0, xend = Comp.1, yend = variable),
               arrow = arrow(length = unit(1/2, "picas"), type='closed'),
               color = "black") +
  geom_vline(aes(xintercept = 0), linetype  = "solid") +
  ggtitle("PC1") +
  labs(x = "", y = " ") +
  theme_bw (base_size = 8) 
 
 
pc2 <- 
  ggplot(mydata, aes(x = Comp.2, y = variable)) +
  xlim(-0.75, 0.75) +
  geom_segment(aes(x = 0, xend = Comp.2, yend = variable),
               arrow = arrow(length = unit(1/2, "picas"), type='closed'),
               color = "black") +
  geom_vline(aes(xintercept = 0), linetype  = "solid") +
  ggtitle("PC2") +
  labs(x = "", y = " ") +
  theme_bw(base_size = 8)
  
 
pc3 <- 
  ggplot(mydata, aes(x = Comp.3, y = variable)) +
  xlim(-0.75, 0.75) +
  geom_segment(aes(x = 0, xend = Comp.3, yend = variable),
               arrow = arrow(length = unit(1/2, "picas"), type='closed'),
               color = "black") +
  geom_vline(aes(xintercept = 0), linetype  = "solid") +
  ggtitle("PC3") +
  labs(x = "", y = " ") +
  theme_bw(base_size = 8)


# provide three seperate panels for the pc loadings
# ggsave(pc1, height = 1.2 , width = 3.25, filename = here("figures/loadings-pca-plots-ALL-pc1.png"))  
# ggsave(pc2, height = 1.2 , width = 3.25, filename = here("figures/loadings-pca-plots-ALL-pc2.png"))  
# ggsave(pc3, height = 1.2 , width = 3.25, filename = here("figures/loadings-pca-plots-ALL-pc3.png"))  


# provide combined panel  
# loadingsall <- ggarrange(pc1, pc2, pc3, height = 7 , width = 7, common.legend = TRUE, #legend = "top", ncol = 1, nrow = 3)
# loadingsall

#doesn't work. not sure why
#ggsave(loadingsall, height = 2 , width = 7, filename = here("figures/loadings-pca-plots-ALL.png"))  
 


```
  

FIGURE SI 4

Plot PC plot with convex hulls for all animals (including juveniles) (n = 91).
```{r, echo = FALSE}
pc12 <-
  ggplot(pog2, aes(x = Comp.1, y = Comp.2, 
                 colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_point(alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
    scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (34.10%)", y = "PC2 (26.73%)") +
  theme(legend.position = "none")

pc13 <-
  ggplot(pog2, aes(x = Comp.1, y = Comp.3, 
                 colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_point(alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
    scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (34.10%)", y = "PC3 (21.79%)") +
  theme(legend.position = "none")

pc23 <-
  ggplot(pog2, aes(x = Comp.2, y = Comp.3, 
                 colour = Genotype, fill = Genotype, shape = Genotype)) +
  geom_point(alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
    scale_shape_manual(values=c(23, 21, 22)) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC2 (26.73%)", y = "PC3 (21.79%)")

all <- ggarrange(pc12, pc13, pc23, height = 7 , width = 7, common.legend = TRUE, legend = "top")
all

#ggsave(all, height = 7 , width = 7, filename = here("figures/pca-plots-ALL.png"))
```





Plot PC plot with convex hulls and scaled points according to bite force for all animals (including juveniles) (n = 91): 
```{r, echo = FALSE}
pc12b <-
  ggplot(pog2, aes(x = Comp.1, y = Comp.2, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = BiteForce), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (34.10%)", y = "PC2 (26.73%)") +
  theme(legend.position = "none")

pc13b <-
  ggplot(pog2, aes(x = Comp.1, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = BiteForce), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (34.10%)", y = "PC3 (21.79%)") +
  theme(legend.position = "none")

pc23b <-
  ggplot(pog2, aes(x = Comp.2, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = BiteForce), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC2 (26.73%)", y = "PC3 (21.79%)")

all <- ggarrange(pc12b, pc13b, pc23b, common.legend = TRUE, legend = "top")
all

#ggsave(all, filename = here("figures/pca-plots-ALL-with_BF.png"))
```


Plot PC plot with convex hulls and scaled points according to SVL for all animals (including juveniles) (n = 91).: 
```{r, echo = FALSE}
pc12b <-
  ggplot(pog2, aes(x = Comp.1, y = Comp.2, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = SVL), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (34.10%)", y = "PC2 (26.73%)") +
  theme(legend.position = "none")

pc13b <-
  ggplot(pog2, aes(x = Comp.1, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = SVL), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (34.10%)", y = "PC3 (21.79%)") +
  theme(legend.position = "none")

pc23b <-
  ggplot(pog2, aes(x = Comp.2, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = SVL), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC2 (26.73%)", y = "PC3 (21.79%)")

all <- ggarrange(pc12b, pc13b, pc23b, common.legend = TRUE, legend = "top")
all

#ggsave(all, height = 12 , width = 7, filename = here("figures/pca-plots-withSVL.png"))
```

Plot PC plot with convex hulls and scaled points according to HDqu for all animals (including juveniles) (n = 91).: 
```{r, echo = FALSE}
pc12b <-
  ggplot(pog2, aes(x = Comp.1, y = Comp.2, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = HWqu), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (34.10%)", y = "PC2 (26.73%)") +
  theme(legend.position = "none")

pc13b <-
  ggplot(pog2, aes(x = Comp.1, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = HWqu), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (34.10%)", y = "PC3 (21.79%)") +
  theme(legend.position = "none")

pc23b <-
  ggplot(pog2, aes(x = Comp.2, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = HWqu), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC2 (26.73%)", y = "PC3 (21.79%)")

all <- ggarrange(pc12b, pc13b, pc23b, common.legend = TRUE, legend = "top")
all

#ggsave(all, height = 12 , width = 7, filename = here("figures/pca-plots-withHWqu.png"))
```



Plot PC plot with convex hulls and scaled points according to HDmt for all animals (including juveniles) (n = 91).: 
```{r, echo = FALSE}
pc12b <-
  ggplot(pog2, aes(x = Comp.1, y = Comp.2, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = HDmt), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (34.10%)", y = "PC2 (26.73%)") +
  theme(legend.position = "none")

pc13b <-
  ggplot(pog2, aes(x = Comp.1, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = HDmt), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (34.10%)", y = "PC3 (21.79%)") +
  theme(legend.position = "none")

pc23b <-
  ggplot(pog2, aes(x = Comp.2, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = HDmt), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC2 (26.73%)", y = "PC3 (21.79%)")

all <- ggarrange(pc12b, pc13b, pc23b, common.legend = TRUE, legend = "top")
all

#ggsave(all, height = 12 , width = 7, filename = here("figures/pca-plots-withHDmt.png"))
```


~~ ADULTS ONLY ~~

Create a dataset with only adults (n = 39).

```{r}
pogA <- 
  pog %>%
  filter(Age == "Adult")
```


PCA for adults

```{r, echo = FALSE, warning = FALSE, message = FALSE}
#First process the data so we can use it...

# Perform a principal components analysis on log shape variables
pca_head_adults <- princomp(as.matrix(select(pogA, LSR_HLroret:LSR_HDmt)),
                     scores = TRUE, cor = TRUE)

# Recombine this with the full dataset
# First make a dataset of just the PC scores and label the rows with ID
# taken from the dataset as these are in the same order
pca_head_ds_adults <- data.frame(pca_head$scores, ID = pog$ID)

# Next join this on "ID" to the original data
pog3 <- left_join(pogA, pca_head_ds_adults, by = "ID")
```

PCA outputs for adults 
```{r}
loadings(pca_head_adults)
summary(pca_head_adults)
```

FIGURE SI 1

Plot PC plot with convex hulls for adults (n = 39):
```{r, echo = FALSE}
Apc12 <-
  ggplot(pog3, aes(x = Comp.1, y = Comp.2, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) + 
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (36.56%)", y = "PC2 (31.69%)") +
  theme(legend.position = "none")

Apc13 <-
  ggplot(pog3, aes(x = Comp.1, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (36.56%)", y = "PC3 (21.19%)") +
  theme(legend.position = "none")

Apc23 <-
  ggplot(pog3, aes(x = Comp.2, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw(base_size = 14) +
  labs(x = "PC2 (31.69%)", y = "PC3 (21.19%)")

all <- ggarrange(Apc12, Apc13, Apc23, common.legend = TRUE, legend = "top")
all

#ggsave(all, height = 7 , width = 7, filename = here("figures/pca-plots-adults.png"))
```


FIGURE SI 2

Plot PC plot with convex hulls and scaled points according to bite force for adults (n = 39): 
```{r, echo = FALSE}
Apc12b <-
  ggplot(pog3, aes(x = Comp.1, y = Comp.2, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = BiteForce), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (36.56%)", y = "PC2 (31.69%)") +
  theme(legend.position = "none")

Apc13b <-
  ggplot(pog3, aes(x = Comp.1, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = BiteForce), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (36.69%)", y = "PC3 (21.19%)") +
  theme(legend.position = "none")

Apc23b <-
  ggplot(pog3, aes(x = Comp.2, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = BiteForce), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC2 (31.69%)", y = "PC3 (21.19%)")

all <- ggarrange(Apc12b, Apc13b, Apc23b, common.legend = TRUE, legend = "top")
all

#ggsave(all, height = 7 , width = 7, filename = here ("figures/pca-plots-Adults-with_BF.png"))
```


FIGURE SI 3

Plot PC plot with convex hulls and scaled points according to SVL for adults (n = 39): 
```{r, echo = FALSE}
Apc12b <-
  ggplot(pog3, aes(x = Comp.1, y = Comp.2, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = SVL), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (36.69%)", y = "PC2 (31.69%)") +
  theme(legend.position = "none")

Apc13b <-
  ggplot(pog3, aes(x = Comp.1, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = SVL), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (36.69%)", y = "PC3 (21.19%)") +
  theme(legend.position = "none")

Apc23b <-
  ggplot(pog3, aes(x = Comp.2, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = SVL), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC2 (31.69%)", y = "PC3 (21.19%)")

all <- ggarrange(Apc12b, Apc13b, Apc23b, common.legend = TRUE, legend = "top")
all

#ggsave(all, height = 7 , width = 7, filename = here ("figures/pca-plots-Adults-with-SVL.png"))

```



Plot PC plot with convex hulls and scaled points according to HDmt for adults (n = 39): 

```{r, echo = FALSE}
Apc12b <-
  ggplot(pog3, aes(x = Comp.1, y = Comp.2, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = HDmt), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (36.69%)", y = "PC2 (31.69%)") +
  theme(legend.position = "none")

Apc13b <-
  ggplot(pog3, aes(x = Comp.1, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = HDmt), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC1 (36.69%)", y = "PC3 (21.19%)") +
  theme(legend.position = "none")

Apc23b <-
  ggplot(pog3, aes(x = Comp.2, y = Comp.3, 
                 colour = Genotype, fill = Genotype)) +
  geom_point(aes(size = HDmt), shape = 19 , alpha = 0.8) +
  geom_convexhull(alpha = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  labs(x = "PC2 (31.69%)", y = "PC3 (21.19%)")

all <- ggarrange(Apc12b, Apc13b, Apc23b, common.legend = TRUE, legend = "top")
all

#ggsave(all, height = 7 , width = 7, filename = here ("figures/pca-plots-Adults-with-HDmt.png"))
```