---
title: "07-temperature"
author: "Marc Jones"
date: "September 5, 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the libraries
```{r, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(ggfortify)
library(here)
```


Read in the dataset, and create a dataset with only adults
```{r}
pog <- read_csv(here("data/pogona-data.csv"))

pogA <- 
  pog %>%
  filter(Age == "Adult")
```

Relevel the genotype variable so everything is compared to ZZf as that's the key comparison...

```{r}
pogA <-
  pogA %>%
  mutate(Genotype = relevel(as.factor(Genotype), ref = "ZZf"))
```



1a Temperature table for whole sample

```{r, warning = FALSE}

pogtempsum <- 
  pog %>%
  group_by(Genotype) %>%
  summarize_at(vars(IncTemp, Temp), 
               list(mean = ~ mean(., na.rm = TRUE), SE = se)) %>%
  mutate_if(is.numeric, round, digits = 2)

kable(pogtempsum)

#write_csv(pogtempsum, path = here("outputs/temperature-all.csv"))

```


1b Temperature table for adults

```{r, warning = FALSE}

pogtempsum <- 
  pogA %>%
  group_by(Genotype) %>%
  summarize_at(vars(IncTemp, Temp), 
               list(mean = ~ mean(., na.rm = TRUE), SE = se)) %>%
  mutate_if(is.numeric, round, digits = 2)

kable(pogtempsum)

#write_csv(pogtempsum, path = here("outputs/temperature-ad.csv"))

```

1c Temperature table for juveniles

```{r, warning = FALSE}

pogtempsum <- 
  pogJ %>%
  group_by(Genotype) %>%
  summarize_at(vars(IncTemp, Temp), 
               list(mean = ~ mean(., na.rm = TRUE), SE = se)) %>%
  mutate_if(is.numeric, round, digits = 2)

kable(pogtempsum)

#write_csv(pogtempsum, path = here("outputs/temperature-juv.csv"))
```



2. Temperature plots 

For adults

```{r, warning = FALSE}
tempplot <-
  ggplot(pogA, aes(x = Genotype, y = Temp, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogtempsum
  geom_point(data = pogtempsum, aes(y = Temp_mean), size = 2) +
  geom_errorbar(data = pogtempsum, aes(y = Temp_mean,
                                    ymin = Temp_mean - Temp_SE,
                                    ymax = Temp_mean + Temp_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(20, 40) +
  ylab(expression('Temperature (°C)'))

inctempplot <-
  ggplot(pogA, aes(x = Genotype, y = IncTemp, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from ppogtempsum
  geom_point(data = pogtempsum, aes(y = IncTemp_mean), size = 2) +
  geom_errorbar(data = pogtempsum, aes(y = IncTemp_mean,
                                    ymin = IncTemp_mean - IncTemp_SE,
                                    ymax = IncTemp_mean + IncTemp_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(20, 40) +
  ylab(expression('Incubation temperature (°C)'))


(tempplot + inctempplot)

#ggsave(here("figures/genotype-means-temp-inctemp.png"), height = 5)

```

For juveniles

```{r, warning = FALSE}
tempplot <-
  ggplot(pogJ, aes(x = Genotype, y = Temp, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogtempsum
  geom_point(data = pogtempsum, aes(y = Temp_mean), size = 2) +
  geom_errorbar(data = pogtempsum, aes(y = Temp_mean,
                                    ymin = Temp_mean - Temp_SE,
                                    ymax = Temp_mean + Temp_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(20, 40) +
  ylab(expression('Temperature (°C)'))

inctempplot <-
  ggplot(pogJ, aes(x = Genotype, y = IncTemp, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from ppogtempsum
  geom_point(data = pogtempsum, aes(y = IncTemp_mean), size = 2) +
  geom_errorbar(data = pogtempsum, aes(y = IncTemp_mean,
                                    ymin = IncTemp_mean - IncTemp_SE,
                                    ymax = IncTemp_mean + IncTemp_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(20, 40) +
  ylab(expression('Incubation temperature (°C)'))


(tempplot + inctempplot)

#ggsave(here("figures/genotype-means-temp-inctemp.png"), height = 5)

```

3. 
This code creates a plot of BF vs log(SVL) colour coded according to temperature.
Draft of Figure S7.

```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(BiteForce), 
                colour = Temp, fill = Temp)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2) +
  #scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  #scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(Bite Force) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/temp-bf.png"))
```

4.
This code creates a plot of BF vs log(SVL) colour coded according to temperature divided by genotype. Actual Figure S7.

```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(BiteForce), 
                colour = Temp, fill = Temp, shape = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE, size = 2) +
  #geom_smooth(method = lm, se = TRUE, alpha = 0.2, colour = "black", size = 0.5) +
  scale_colour_gradientn(name = "Temperature (°C)", colours = heat.colors(10)) +
  scale_fill_gradientn(name = "Temperature (°C)", colours = heat.colors(10)) +
  scale_shape_manual(values=c(23, 21, 22)) +
  theme_bw(base_size = 14) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(Bite Force) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/temp-genotype-bf-shapes.png"))
```



5.
This code creates a plot of BF vs log(SVL) colour coded according to inclubation temperature divided by genotype.
Actual Figure S8.


```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(BiteForce), 
                colour = IncTemp, fill = IncTemp, shape = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE, size = 2) +
  #geom_smooth(method = lm, se = TRUE, alpha = 0.2, colour = "black", size = 0.5) +
  scale_colour_gradientn(name = "Incubation Temp (°C)", colours = heat.colors(10)) +
  scale_fill_gradientn(name = "Incubation Temp (°C)", colours = heat.colors(10)) +
  scale_shape_manual(values=c(23, 21, 22)) +
  theme_bw(base_size = 14) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(Bite Force) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/inctemp-genotype-bf-shapes.png"))
```


6.
Summary statistics.

Set up an empty output file
```{r}
output <- data.frame(array(dim = c(2, 20)))
names(output) <- c("response", "predictor", "df", 
                   "F_pred", "p_pred", 
                   "F_gen", "p_gen", 
                   "F_int", "p_int", 
                   "slope_ZZF", "slope_se_ZZF",
                   "slopediff_ZWF", "se_ZWF",
                   "t_ZWF", "p_ZWF",
                   "slopediff_ZZM", "se_ZZM",
                   "t_ZZM", "p_ZZM",
                   "adj.r2")

# Short function to help write output files
write_output <- function(model, response, predictor, n){
  
  x <- tidy(anova(model))
  y <- summary(model)
    
  output[n, "response"] <- response
  output[n, "predictor"] <- predictor
  output[n, "df"] <- x$df[4]
    output[n, "F_pred"] <- x$statistic[1]
    output[n, "p_pred"] <- x$p.value[1]
    output[n, "F_gen"] <- x$statistic[2]
    output[n, "p_gen"] <- x$p.value[2]
    output[n, "F_int"] <- x$statistic[3]
    output[n, "p_int"] <- x$p.value[3]
    output[n, "slope_ZZF"] <- y$coefficients[2]
    output[n, "slope_se_ZZF"] <- y$coefficients[8]
    output[n, "slopediff_ZWF"] <- y$coefficients[5]
    output[n, "se_ZWF"] <- y$coefficients[11]
    output[n, "t_ZWF"] <- y$coefficients[17]
    output[n, "p_ZWF"] <- y$coefficients[23]
    output[n, "slopediff_ZZM"] <- y$coefficients[6]
    output[n, "se_ZZM"] <- y$coefficients[12]
    output[n, "t_ZZM"] <- y$coefficients[18]
    output[n, "p_ZZM"] <- y$coefficients[24]
    output[n, "adj.r2"] <- y$adj.r.squared
    return(output)
}
```

7.
Temp
```{r}
# Fit the model
t1 <- lm(log(BiteForce) ~ log(Temp) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(t1)
tidy(anova(t1))
summary(t1)

output <- write_output(t1, "log(Bite Force)", "log(temp)", 1)
```




8.
Inctemp
```{r}
# Fit the model
t2 <- lm(log(BiteForce) ~ log(IncTemp) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(t2)
tidy(anova(t2))
summary(t2)

output <- write_output(t2, "log(Bite Force)", "log(IncTemp)", 2)
```



Write all outputs to file
```{r}
write_csv(output, path = here("outputs/bite-force-temp-results-h.csv"))
```







9. 
An ANOVA for all animals, adults, and juveniles.


Write a quick standard error function
```{r}
se <- function(x){
  sqrt(var(x)/length(x))
}
```



Read in the dataset, and create a dataset with only adults
```{r}
pog <- read_csv(here("data/pogona-data.csv"))

pogA <- 
  pog %>%
  filter(Age == "Adult")
```

Read in the dataset, and create a dataset with only juveniles
```{r}
pog <- read_csv(here("data/pogona-data.csv"))

pogJ <- 
  pog %>%
  filter(Age == "Juvenile")
```



Relevel the genotype variable so everything is compared to ZZf as that's the key comparison...

```{r}
pogA <-
  pogA %>%
  mutate(Genotype = relevel(as.factor(Genotype), ref = "ZZf"))
```


Relevel the genotype variable so everything is compared to ZZf as that's the key comparison...

```{r}
pogJ <-
  pogJ %>%
  mutate(Genotype = relevel(as.factor(Genotype), ref = "ZZf"))
```

Fit ANOVAs to look at differences in temperature between the three genotypes.

First to make the models easier to interpret, relevel the genotype variable so everything is compared to ZZf as that's the key comparison...

Next set up an empty output file
```{r}
output <- data.frame(array(dim = c(6, 15)))
names(output) <- c("response", "df1", "df2", "F", "p", 
                  "mean_ZZf", "se_ZZf",
                  "meandiff_ZWf", "se_ZWf", "t_ZWf", "p_ZWf",
                  "meandiff_ZZm", "se_ZZm", "t_ZZm", "p_ZZm")

# Short function to help write output files
write_output <- function(model, response, n){
  
  x <- tidy(anova(model))
  y <- summary(model)
    
  output[n, "response"] <- response
  output[n, "df1"] <- x$df[1]
    output[n, "df2"] <- x$df[2]
    output[n, "F"] <- x$statistic[1]
    output[n, "p"] <- x$p.value[1]
    output[n, "mean_ZZf"] <- y$coefficients[1]
    output[n, "se_ZZf"] <- y$coefficients[4]
    output[n, "meandiff_ZWf"] <- y$coefficients[2]
    output[n, "se_ZWf"] <- y$coefficients[5]
    output[n, "t_ZWf"] <- y$coefficients[8]
    output[n, "p_ZWf"] <- y$coefficients[11]
    output[n, "meandiff_ZZm"] <- y$coefficients[3]
    output[n, "se_ZZm"] <- y$coefficients[5]
    output[n, "t_ZZm"] <- y$coefficients[9]
    output[n, "p_ZZm"] <- y$coefficients[12]
    
    return(output)
}
```


Surface temperature measured using a temperature gun between trial 1 and trial 2
(Temp)

```{r}
# Fit the model
m1 <- lm(log(Temp) ~ Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m10)
tidy(anova(m10))
summary(m10)

output <- write_output(m1, "log(Temp)", 1)
```


Adult surface temperature measured using a temperature gun between trial 1 and trial 2
(Temp)

```{r}
# Fit the model
m2 <- lm(log(Temp) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m10)
tidy(anova(m10))
summary(m10)

output <- write_output(m2, "log(Temp)", 2)
```

Juvenile surface temperature measured using a temperature gun between trial 1 and trial 2
(Temp)

```{r}
# Fit the model
m3 <- lm(log(Temp) ~ Genotype, data = pogJ)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m10)
tidy(anova(m10))
summary(m10)

output <- write_output(m3, "log(Temp)", 3)
```


Egg incubation temperature
(IncTemp)
```{r}
# Fit the model
m4 <- lm(log(IncTemp) ~ Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m11)
tidy(anova(m11))
summary(m11)

output <- write_output(m4, "log(IncTemp)", 4)
```

Adult egg incubation temperature
(IncTemp)
```{r}
# Fit the model
m5 <- lm(log(IncTemp) ~ Genotype, data = pogA)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m11)
tidy(anova(m11))
summary(m11)

output <- write_output(m5, "log(IncTemp)", 5)
```


Adult egg incubation temperature
(IncTemp)
```{r}
# Fit the model
m6 <- lm(log(IncTemp) ~ Genotype, data = pogJ)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(m11)
tidy(anova(m11))
summary(m11)

output <- write_output(m6, "log(IncTemp)", 6)
```


Write all outputs to file
```{r}
write_csv(output, path = here("outputs/temp-ANOVA-results-e.csv"))
```














