---
title: "07-temperature"
author: "Marc Jones"
date: "September 5, 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the libraries
```{r, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(ggfortify)
library(here)
```

Read in the dataset, relevel the genotype variable so everything is compared to ZZf as that's the key comparison.

```{r}
pog <- read_csv(here("data/pogona-data.csv"))

pog <-
  pog %>%
  mutate(Genotype = relevel(as.factor(Genotype), ref = "ZZF"))
```



1. Temperature table (Table S5)

```{r, warning = FALSE}

pogtempsum <- 
  pogA %>%
  group_by(Genotype) %>%
  summarize_at(vars(IncTemp, Temp), 
               list(mean = ~ mean(., na.rm = TRUE), SE = se)) %>%
  mutate_if(is.numeric, round, digits = 2)

kable(pogtempsum)

#write_csv(pogtempsum, path = here("outputs/temperature.csv"))

```

2. Temperature plots 
(Figure S6)

```{r, warning = FALSE}
tempplot <-
  ggplot(pogA, aes(x = Genotype, y = Temp, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from pogtempsum
  geom_point(data = pogtempsum, aes(y = Temp_mean), size = 2) +
  geom_errorbar(data = pogtempsum, aes(y = Temp_mean,
                                    ymin = Temp_mean - Temp_SE,
                                    ymax = Temp_mean + Temp_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(20, 40) +
  ylab(expression('Temperature (°C)'))

inctempplot <-
  ggplot(pogA, aes(x = Genotype, y = IncTemp, 
                colour = Genotype, fill = Genotype)) +
  geom_jitter(alpha = 0.4, na.rm = TRUE, width = 0.2) +
  scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  # Add mean and se from ppogtempsum
  geom_point(data = pogtempsum, aes(y = IncTemp_mean), size = 2) +
  geom_errorbar(data = pogtempsum, aes(y = IncTemp_mean,
                                    ymin = IncTemp_mean - IncTemp_SE,
                                    ymax = IncTemp_mean + IncTemp_SE),
                width = 0.1) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  ylim(20, 40) +
  ylab(expression('Incubation temperature (°C)'))


(tempplot + inctempplot)

#ggsave(here("figures/genotype-means-temp-inctemp.png"), height = 5)

```

3. 
This code creates a plot of BF vs log(SVL) colour coded according to temperature.
Draft of Figure S7.

```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(BiteForce), 
                colour = Temp, fill = Temp)) +
  geom_point(alpha = 0.8, na.rm = TRUE) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2) +
  #scale_colour_manual(values = c("orange1", "red", "cornflowerblue")) +
  #scale_fill_manual(values = c("orange1", "red", "cornflowerblue")) +
  theme_bw(base_size = 14) +
  #theme(legend.title = element_blank()) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(Bite Force) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/temp-bf.png"))
```

4.
This code creates a plot of BF vs log(SVL) colour coded according to temperature divided by genotype. Actual Figure S7.

```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(BiteForce), 
                colour = Temp, fill = Temp, shape = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE, size = 2) +
  #geom_smooth(method = lm, se = TRUE, alpha = 0.2, colour = "black", size = 0.5) +
  scale_colour_gradientn(name = "Temperature (°C)", colours = heat.colors(10)) +
  scale_fill_gradientn(name = "Temperature (°C)", colours = heat.colors(10)) +
  scale_shape_manual(values=c(23, 21, 22)) +
  theme_bw(base_size = 14) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(Bite Force) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/temp-genotype-bf-shapes.png"))
```



5.
This code creates a plot of BF vs log(SVL) colour coded according to inclubation temperature divided by genotype.
Actual Figure S8.


```{r, message = FALSE}
# All individuals
ggplot(pog, aes(x = log(SVL), y = log(BiteForce), 
                colour = IncTemp, fill = IncTemp, shape = Genotype)) +
  geom_point(alpha = 0.8, na.rm = TRUE, size = 2) +
  #geom_smooth(method = lm, se = TRUE, alpha = 0.2, colour = "black", size = 0.5) +
  scale_colour_gradientn(name = "Incubation Temp (°C)", colours = heat.colors(10)) +
  scale_fill_gradientn(name = "Incubation Temp (°C)", colours = heat.colors(10)) +
  scale_shape_manual(values=c(23, 21, 22)) +
  theme_bw(base_size = 14) +
  ylim(0, 5)+
  xlim(4, 5.5) +
  ylab("log(Bite Force) (N)") +
  xlab("log(SVL) (mm)")

#ggsave(here("figures/inctemp-genotype-bf-shapes.png"))
```


6.
Summary statistics.

Set up an empty output file
```{r}
output <- data.frame(array(dim = c(2, 20)))
names(output) <- c("response", "predictor", "df", 
                   "F_pred", "p_pred", 
                   "F_gen", "p_gen", 
                   "F_int", "p_int", 
                   "slope_ZZF", "slope_se_ZZF",
                   "slopediff_ZWF", "se_ZWF",
                   "t_ZWF", "p_ZWF",
                   "slopediff_ZZM", "se_ZZM",
                   "t_ZZM", "p_ZZM",
                   "adj.r2")

# Short function to help write output files
write_output <- function(model, response, predictor, n){
  
  x <- tidy(anova(model))
  y <- summary(model)
    
  output[n, "response"] <- response
  output[n, "predictor"] <- predictor
  output[n, "df"] <- x$df[4]
    output[n, "F_pred"] <- x$statistic[1]
    output[n, "p_pred"] <- x$p.value[1]
    output[n, "F_gen"] <- x$statistic[2]
    output[n, "p_gen"] <- x$p.value[2]
    output[n, "F_int"] <- x$statistic[3]
    output[n, "p_int"] <- x$p.value[3]
    output[n, "slope_ZZF"] <- y$coefficients[2]
    output[n, "slope_se_ZZF"] <- y$coefficients[8]
    output[n, "slopediff_ZWF"] <- y$coefficients[5]
    output[n, "se_ZWF"] <- y$coefficients[11]
    output[n, "t_ZWF"] <- y$coefficients[17]
    output[n, "p_ZWF"] <- y$coefficients[23]
    output[n, "slopediff_ZZM"] <- y$coefficients[6]
    output[n, "se_ZZM"] <- y$coefficients[12]
    output[n, "t_ZZM"] <- y$coefficients[18]
    output[n, "p_ZZM"] <- y$coefficients[24]
    output[n, "adj.r2"] <- y$adj.r.squared
    return(output)
}
```

Temp
```{r}
# Fit the model
t1 <- lm(log(BiteForce) ~ log(Temp) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(t1)
tidy(anova(t1))
summary(t1)

#output <- write_output(t1, "log(Bite Force)", "log(temp)", 1)
```

Inctemp
```{r}
# Fit the model
t2 <- lm(log(BiteForce) ~ log(IncTemp) * Genotype, data = pog)
# Look at the model diagnostics (these all look fine so ANOVA is ok)
autoplot(t2)
tidy(anova(t2))
summary(t2)

#output <- write_output(t2, "log(Bite Force)", "log(IncTemp)", 2)
```


Write all outputs to file
```{r}
#write_csv(output, path = here("outputs/bite-force-temp-results.csv"))
```